A thermodynamic view of dusty protoplanetary disks

### 2017-06-29 paper.tex
minor tweaks to abstract

\begin{abstract}
  Small solids embedded in gaseous protoplanetary disks are subject
  to strong dust-gas friction. Consequently, tightly-coupled dust
  particles almost follows the gas flow. 
  This is analogous to entropy conservation
  following an ideal gas.   
  Motivated by this similarity, we develop a thermodynamic 
  framework to study dusty gas dynamics in protoplanetary disks. We  
  show that an isothermal dusty gas behaves like an  
  adiabatic pure gas; and that finite dust-gas coupling may be
  regarded as an effective 
  heating/cooling. We exploit this correspondence 
  to deduce that   
\begin{inparaenum}[1)] 
\item 
 perfectly coupled, thin dust layers cannot cause axisymmetric 
 instabilties; 
\item 
  the streaming instability neccessarily involves a gas pressure
  response 
  that lags behind dust density; 
%  the streaming instability neccessarily involves a phase lag between
%  gas pressure and dust densities; 
\item dust-loading introduce buoyancy forces that generally 
  stabilizes the vertical shear 
  instability associated with global radial temperature gradients. 
\end{inparaenum}  
%growth rates, and 
%preferentially stabilize modes with small radial 
%wavelengths, as well as `surface modes' concentrated near the
%disk boundaries. 
%These results can be understood in terms of 
%buoyancy forces in an isothermal dusty disk. 
%We show that vertical shear arising from
%dust settling does not lead to axisymmetric instability. 
We also discuss %future applications of this thermodynamic framework in
%finding 
potential dusty analogs of other hydrodynamic 
processes (e.g. Rossby wave instability, convective overstability, and 
zombie vortices), in simulating dusty protoplanetary disks
using purely hydrodynamic codes, as well as possible generalizations
of the dusty/adiabatic gas analogy.

>>>>

\begin{abstract}
  Small solids embedded in gaseous protoplanetary disks are subject
  to strong dust-gas friction. Consequently, tightly-coupled dust
  particles almost follow the gas flow.
  %AY: Edited b/c "This is" was too vague
  This near conservation of dust-to-gas ratio along streamlines
  is analogous to the near conservation of entropy along flows of (dust-free) gas with weak heating and cooling.   
  We develop this thermodynamic analogy into a
  framework to study dusty gas dynamics in protoplanetary disks. We  
  show that an isothermal dusty gas behaves like an  
  adiabatic pure gas; and that finite dust-gas coupling may be
  regarded as an effective 
  heating/cooling. We exploit this correspondence 
  to deduce that   
\begin{inparaenum}[1)] 
\item 
 perfectly coupled, thin dust layers cannot cause axisymmetric 
 instabilties; 
\item 
  the streaming instability neccessarily involves a gas pressure
  response 
  that lags behind dust density; 
%  the streaming instability neccessarily involves a phase lag between
%  gas pressure and dust densities; 
\item dust-loading introduces buoyancy forces that generally 
  stabilizes the vertical shear 
  instability associated with global radial temperature gradients. 
\end{inparaenum}  
%growth rates, and 
%preferentially stabilize modes with small radial 
%wavelengths, as well as `surface modes' concentrated near the
%disk boundaries. 
%These results can be understood in terms of 
%buoyancy forces in an isothermal dusty disk. 
%We show that vertical shear arising from
%dust settling does not lead to axisymmetric instability. 
We also discuss %future applications of this thermodynamic framework in
%finding 
 dusty analogs of other hydrodynamic 
processes (e.g. Rossby wave instability, convective overstability, and 
zombie vortices), how to simulate dusty protoplanetary disks
with minor tweaks to purely hydrodynamic codes, as well as possible generalizations
of our analogy between dust-laden and nearly adiabatic flows.
